---
title: "Citation searching with paperfetcher"
output: html_notebook
---

This [R Markdown](http://rmarkdown.rstudio.com) Notebook demonstrates how to use the paperfetcher Python package to perform forward and backward citation searching in R through the [reticulate](https://https://rstudio.github.io/reticulate/) interface to Python.

To execute a code chunk in RStudio, click the *Run* button within the chunk or place your cursor inside it and press *Cmd+Shift+Enter*.

# Setup

We first need to run a couple of lines of code to install reticulate, Python, and paperfetcher.

```{r}
# Install reticulate if not already installed
if(!require("reticulate")){
  install.packages("reticulate")
  library(reticulate)
}

# Install Python if not already installed, and create a new virtualenv for paperfetcher
# (Uncomment the lines below to run code)
#install_python("3.7:latest")
#virtualenv_create("paperfetcher", version="3.7:latest")
#use_virtualenv("paperfetcher")

# Install paperfetcher
# (Uncomment the lines below to run code)
#py_install("paperfetcher", envname="paperfetcher")

# Import the paperfetcher package
paperfetcher <- import("paperfetcher")
```

# Snowballing backwards (also called backward reference chasing, backward reference search, or backward citation search) with Crossref

Backward reference chasing involves retrieving all articles which are referenced (cited) by a set of starting articles.

Let's fetch all the references from two papers with DOIs:

- 10.1021/acs.jpcb.1c02191
- 10.1073/10.1080/07448481.2022.2059376

using the Crossref service.

First, we create a search object, and initialize it with a list of strings, each string being a DOI:

```{r}
search <- paperfetcher$snowballsearch$CrossrefBackwardReferenceSearch(list("10.1021/acs.jpcb.1c02191", "10.1080/07448481.2022.2059376"))
search()
```

How many articles did our search return?

```{r}
py_len(search)
```

# Extracting data from the search results

Just as we did for handsearching, we can get a Dataset of DOIs from the search results:

```{r}
doi_ds <- search$get_DOIDataset()
```

We can display this as a DataFrame:

```{r}
doi_ds$to_df()
```
Or save it to a text file:

```{r}
doi_ds$save_txt("out/snowball_back.txt")
```

We can also convert it to RIS format:

```{r}
ris_ds <- search$get_RISDataset()
```

And save it to an RIS file:

```{r}
ris_ds$save_ris("out/snowball_back.ris")
```

# Snowballing backwards with COCI

We can also perform backward snowballing with COCI, the OpenCitations Index of Crossref DOI-to-DOI citations.

The syntax is similar to that of Crossref:

```{r}
search <- paperfetcher$snowballsearch$COCIBackwardReferenceSearch(list("10.1021/acs.jpcb.1c02191", "10.1080/07448481.2022.2059376"))
search()
doi_ds <- search$get_DOIDataset()
doi_ds$to_df()
```

# Snowballing forwards (also called forward citation chasing or forward citation search) with COCI

Forward citation chasing involves retrieving all articles which cite a set of starting articles.

Let's fetch all the citations of two papers with DOIs:

- 10.1021/acs.jpcb.1c02191
- 10.1073/10.1080/07448481.2022.2059376

using the COCI service. We cannot use the Crossref service for this task.

The syntax is similar to that of backward search:

```{r}
search <- paperfetcher$snowballsearch$COCIForwardCitationSearch(list("10.1021/acs.jpcb.1c02191", "10.1080/07448481.2022.2059376"))
search()
doi_ds <- search$get_DOIDataset()
doi_ds$to_df()
```

Again, we can save the search results to a text file:

```{r}
doi_ds$save_txt("out/snowball_fwd.txt")
```

Or to an RIS file:

```{r}
ris_ds <- search$get_RISDataset()
ris_ds$save_ris("out/snowball_fwd.ris")
```




